---
title: "Assignment4_part2"
author: "Zirui Liu"
date: "3/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(stringr)
library(foreign)
library(plyr)
library(reshape2)
source("pew.R")
part1 <- raw
source("billboard.R")
```

## Table 4 -> Table 6
### Table 4
```{r table4, ech=T}
part1[1:10, 1:7]
```
For table4, we omit three columns: $75-100k, $100-150k and $>150k. The rows are religion and incomes.

### Table 6
```{r table6, echo=T}
tb6 <- as.tibble(part1)
tb6%>% gather(-religion, key = "income", value = "frequency") %>% arrange(religion) %>% head(n=20)
```
For table6, the rows are religion names, income range and frequency. There are only 20 rows shown.

## Table 7 -> Table 8
### Table 7
```{r table7, echo=T}
raw[c(1:3, 6:10), 1:8]
```
For table7, it only includes top 8 hits for 2000 on Billboard.

### Table 8
```{r table8, echo = T}
table7 <- as.tibble(raw)
table8 <- table7 %>% gather(key= "week", value= "rank", -year, -artist, -track, -time, -date.entered ) %>%
  arrange(artist) %>%
  select(year,artist, time, track, date=date.entered, week, rank) %>%
  filter(!is.na(rank))
head(table8,n=20)
```
For table8, it has date columns. Weeks and ranks are new columns listed for each song. There are only 20 rows shown.

